---
title: "Reproducible Research: Peer Assessment 1"
author: "Andrea Fatutta"
date: "06 april 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup locale and library, unzipping file
```{r loclib}
Sys.setlocale("LC_TIME", "English")
library(lattice)
unzip("activity.zip")
```

###1.Code for reading in the dataset and/or processing the data
```{r lab1}
file <- read.csv("activity.csv", sep = ",", na.strings = "NA")
file$date<-as.Date(file$date,"%Y-%m-%d")
file$steps<- as.numeric(as.character(file$steps))

sumDay <- aggregate(steps~date,file, sum,na.rm=TRUE)
avgInterval <- aggregate(steps~interval,file, mean,na.rm=TRUE)
```


###2.Histogram of the total number of steps taken each day
```{r lab2}
histogram(sumDay$steps, main="Total steps per day", xlab = "Number of steps", breaks = seq(0,22000,by = 2000), col = "Blue" )
```




###3.Mean and median number of steps taken each day
```{r lab3}
median(sumDay$steps)
mean(sumDay$steps)
```


###4.Time series plot of the average number of steps taken
```{r lab4}
xyplot( steps~interval , data = avgInterval, type="a", col="Black", 
        main="Average steps per interval", xlab = "Interval by 5 minutes", ylab = "Number of steps",
        panel = function(x, y, subscripts, ...) {
           panel.xyplot(x, y,subscripts = subscripts, ...)
           panel.abline(v = avgInterval[ avgInterval$steps == max(avgInterval$steps),"interval"], col="Red", lwd= 2)
           }
       )
```

###5.The 5-minute interval that, on average, contains the maximum number of steps
```{r lab5}
paste("Interval with maximum average number of steps:",avgInterval[ avgInterval$steps == max(avgInterval$steps),"interval"])

```

###6.Code to describe and show a strategy for imputing missing data
```{r lab6}
paste("Number of missing steps:", sum(is.na(file$steps)))
paste("Number of missing dates:", sum(is.na(file$date)))
paste("Number of missing interval:", sum(is.na(file$interval)))

print("I will replace missing steps with the average on the interval across all days")

fileFilled <- file

for (i in fileFilled[is.na(fileFilled$steps),"interval"]) 
{
    fileFilled[is.na(fileFilled$steps) & fileFilled$interval == i,"steps"]<- avgInterval[avgInterval$interval==i,"steps"]
}

```

###7.Histogram of the total number of steps taken each day after missing values are imputed
```{r lab7}
sumDayF <- aggregate(steps~date,fileFilled, sum,na.rm=TRUE)
avgIntervalF <- aggregate(steps~interval,fileFilled, mean,na.rm=TRUE)

histogram(sumDayF$steps, main="Total steps per day (filled missing data)", xlab = "Number of steps", breaks = seq(0,22000,by = 2000), col = "Blue" )

median(sumDayF$steps)
mean(sumDayF$steps)
```

###8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

First step, I'll add factor to file and fileFilled data frames, then I draw the plots

```{r lab8}
file$dayType <-factor(sapply(X=weekdays(file$date),FUN = switch,
                             "Monday" = "weekday", 
                             "Tuesday" = "weekday", 
                             "Wednesday" = "weekday", 
                             "Thursday" = "weekday", 
                             "Friday" = "weekday", 
                             "Saturday" = "weekend", 
                             "Sunday" = "weekend", 
                             ""))

fileFilled$dayType <-factor(sapply(X=weekdays(fileFilled$date),FUN = switch,
                             "Monday" = "weekday", 
                             "Tuesday" = "weekday", 
                             "Wednesday" = "weekday", 
                             "Thursday" = "weekday", 
                             "Friday" = "weekday", 
                             "Saturday" = "weekend", 
                             "Sunday" = "weekend", 
                             ""))

avgIntervaldT <- aggregate(steps~interval+dayType,file, mean,na.rm=TRUE)
avgIntervalFdT<- aggregate(steps~interval+dayType,fileFilled, mean,na.rm=TRUE)

xyplot( steps ~ interval| dayType ,data = avgIntervaldT, layout=c(1,2), 
        xlab = "Interval", ylab = "number of steps", type="a", 
        main="Comparison between \n weekdays and weekends - original data")

xyplot( steps ~ interval| dayType ,data = avgIntervalFdT, layout=c(1,2), 
        xlab = "Interval", ylab = "number of steps", type="a",
        main="Comparison between \n weekdays and weekends - filled \"NA\" data")
```





